---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Ryan Wheat"
date: "Insert date here"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
dennys <- dennys
laquinta <- laquinta
```

### Exercise 1

There are 1,643 rows, each representing one Denny's location. There are 6 columns, each representing one piece of info about each respective location: address, city, state, zip, longitude, lattitude.

```{r denny's-dimensions}

#number of rows n columns

nrow(dennys)
ncol(dennys)

#glimpse of rows n columns

glimpse(dennys)
summary(dennys)

```


### Exercise 2

There are 909 rows, each representing one La Quinta hotel location. There are, again, 6 columns, each of which represent one variable for each location: address, city, state, zip, longitude, lattitude. This is the same dataframe structure as the Denny's dataset.

```{r laquinta-dimensions}

#number of rows n columns

nrow(laquinta)
ncol(laquinta)

#what are the variables

glimpse(laquinta)
summarise(laquinta)

```


### Exercise 3

There are no Denny's locations outside the U.S.A. There are 2,295 La Quinta locations in other countries. They are in the following places: Africa, Asia, Australia/Pacific Rim, Carribean, Central America, Europe, Middle East, South America.

### Exercise 4

You could tell whether locations in the La Quinta dataframe are outside the U.S. by filtering based on the state column.

### Exercise 5

There are no Dennys locations outside of the U.S.A.! Woo!

```{r dennys-outside?}

#filtering out Denny's locations that are outside the U.S.

dennys %>%
  filter(!(state %in% states$abbreviation))

```

### Exercise 6

...
```{r dennys-country}

#creating new variable "Country" for Dennys locations

dennys <- dennys %>%
  mutate(country = "United States")
```

### Exercise 7

There are La Quinta locations outside of America in Canada, China, Mexico, New Zealand, Turkey, UAE, Chile, & Colombia.


### Exercise 8

```{r laquinta-country}

#Creating new variable "Country" for La Quinta locations. Note that the countries do not match what is on the La Quinta website (where I got the previous info...)

laquinta <- laquinta %>%
  mutate(country = case_when(
    state %in% states$abbreviation ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state %in% c("QR", "AG", "CH", "NL", "VE", "PU", "SL") ~ "Mexico",
    state ==  "FM" ~ "Honduras"))

#filtering La Quinta locations for only those in U.S.A.

laquinta <- laquinta %>%
  filter(country == "United States")

```


### Exercise 9

```{r locations-by-state}

dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

laquinta %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))

```


Add exercise headings as needed.

